---
title: "Project"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Packages

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(tidymodels)
library(GGally)
library(glmnet)
library(ranger)
options(ranger.num.threads = 16)
library(xgboost)
library(kknn)
library(performance)
```

PROJECT

– clean and explore the data; – study the feasibility of the questions;
– propose different methods to answer the questions and study their
advantages and limitations; – develop visualizations to explain both the
data and the methods’ advantages and limitations; – present the results
in a concise way through a presentation; – present the work in a more
extensive way with a report.

Question setup: We are aliens who are looking at Earth. We have a vague
understanding that the territory is separated by geographical regions,
countries and continents, and each has a choice on how to run its own
territory. We land in a determined region and want to see how to improve
it.

First of all: determine how to evaluate a country: (outcome variables) -
infant mortality rate - maternal mortality rate - life expectancy at
birth - air pollution death rate - crude suicide rates (??) - incidents
of malaria - incidents of tuberculosis - poisoning mortality rate (??) -
neonatal mortality rate - new HIV infections (??) - road traffic
deaths - under 5 mortality rate

determine what a country can change to influence the above outcomes:
(decision variables) - number of doctors - number of dentists - number
of pharmacists - basic drinking water services - least basic sanitation
services - births attended by skilled personnel - basic hand washing -
clean fuel and tech - tobacco at age 15 (advertisement) (maybe not
this) - nursing and midwife numbers - usage of sanitation services

POTENTAIL QUESTIONS TO BE ANSWERED

Q1. Which are the most impactful features (decision variables) when
predicting outcome variables (lin reg) Q2. predict classification
outcome variables based in decision variables Q3. do this for a specific
country/continent (looking at the historic data and all different
features) Q4. (maybe) clustering by Continent based on important outcome
features

Hypothesis: Higher literacy/education levels are associated with better
health outcomes (e.g., lower infant mortality). Hypothesis: Female have
the same life expectancy at birth to males Hypothesis: Female infant
mortality is the same to male infant mortality.

correlations Hypothesis: incidence of diseases (Tuberculosis, Malaria)
correlate with infant mortality rate Hypothesis: medical doctors
correlate with life expectancy at birth ...

classification classify continent based on features (?specific year)

Loading data which might seem useful to answer statistical questions

```{r}

# outcome features
infant_mortality_rate_data = read_csv("data/infantMortalityRate.csv")
maternal_mortality_ratio_data = read_csv("data/maternalMortalityRatio.csv")
life_expectency_at_birth_data = read_csv("data/lifeExpectancyAtBirth.csv")
incidence_of_malaria_data = read_csv("data/incedenceOfMalaria.csv")
incidence_of_tuberculosis_data = read_csv("data/incedenceOfTuberculosis.csv")
crude_sucide_rates = read_csv("data/crudeSuicideRates.csv")
poison_mortality_rate = read_csv("data/mortalityRatePoisoning.csv")
neo_natal_mortality_rate = read_csv("data/neonatalMortalityRate.csv")
new_hiv_infections = read_csv("data/newHivInfections.csv")
road_traffic_deaths = read_csv("data/roadTrafficDeaths.csv")
under_5_mortality_rate = read_csv("data/under5MortalityRate.csv")
# air_pollution_death_rate = read_csv("data/airPollutionDeathRate.csv")   has too many features and is only for 2016


# decision features
medical_doctors_data = read_csv("data/medicalDoctors.csv")
pharmacists_data = read_csv("data/pharmacists.csv")
birth_by_skilled_personel_data = read_csv("data/birthAttendedBySkilledPersonal.csv")
number_of_dentists = read_csv("data/dentists.csv")
drinking_water_services = read_csv("data/basicDrinkingWaterServices.csv")
# least_basic_sanitation_services = read_csv("data/atLeastBasicSanitizationServices.csv")
basic_hand_washing = read_csv("data/basicHandWashing.csv")
clean_fuel_and_teck = read_csv("data/cleanFuelAndTech.csv")
tabacco_age_15 = read_csv("data/tobaccoAge15.csv")
nursing_services = read_csv("data/nursingAndMidwife.csv")
sanitation_services = read_csv("data/safelySanitization.csv")


head(medical_doctors_data, 5)
```

```{r}

# function to remove indicator and rename `First Tooltip` to indicator feature name
replace_inidcator_to_tooltip = function(tibble) {
  df = tibble
  feat_name = tibble$Indicator[1]
  df = subset(df, select = - Indicator)
  names(df)[names(df) == "First Tooltip"] <- feat_name
  
  return(df)
}

```

```{r}

# lowest 1962 year
# highest 2019 year

# creating an empty merged_data so we can add all the features from the datasets

countries = c(
    "Afghanistan", "Albania", "Algeria", "Angola", "Antigua and Barbuda",
    "Argentina", "Armenia", "Australia", "Austria", "Azerbaijan",
    "Bahamas", "Bahrain", "Bangladesh", "Barbados", "Belarus",
    "Belgium", "Belize", "Benin", "Bhutan", "Bolivia (Plurinational State of)",
    "Bosnia and Herzegovina", "Botswana", "Brazil", "Brunei Darussalam",
    "Bulgaria", "Burkina Faso", "Burundi", "Cabo Verde", "Cambodia",
    "Cameroon", "Canada", "Central African Republic", "Chad", "Chile",
    "China", "Colombia", "Comoros", "Congo", "Costa Rica", "Croatia",
    "Cuba", "Cyprus", "Czechia", "Côte d'Ivoire", "Democratic People's Republic of Korea",
    "Democratic Republic of the Congo", "Denmark", "Djibouti",
    "Dominican Republic", "Ecuador", "Egypt", "El Salvador", "Equatorial Guinea",
    "Eritrea", "Estonia", "Eswatini", "Ethiopia", "Fiji", "Finland",
    "France", "Gabon", "Gambia", "Georgia", "Germany", "Ghana",
    "Greece", "Grenada", "Guatemala", "Guinea", "Guinea-Bissau",
    "Guyana", "Haiti", "Honduras", "Hungary", "Iceland", "India",
    "Indonesia", "Iran (Islamic Republic of)", "Iraq", "Ireland",
    "Israel", "Italy", "Jamaica", "Japan", "Jordan", "Kazakhstan",
    "Kenya", "Kiribati", "Kuwait", "Kyrgyzstan", "Lao People's Democratic Republic",
    "Latvia", "Lebanon", "Lesotho", "Liberia", "Libya", "Lithuania",
    "Luxembourg", "Madagascar", "Malawi", "Malaysia", "Maldives",
    "Mali", "Malta", "Mauritania", "Mauritius", "Mexico", "Micronesia (Federated States of)",
    "Mongolia", "Montenegro", "Morocco", "Mozambique", "Myanmar",
    "Namibia", "Nepal", "Netherlands", "New Zealand", "Nicaragua",
    "Niger", "Nigeria", "Norway", "Oman", "Pakistan", "Panama",
    "Papua New Guinea", "Paraguay", "Peru", "Philippines", "Poland",
    "Portugal", "Qatar", "Republic of Korea", "Republic of Moldova",
    "Romania", "Russian Federation", "Rwanda", "Saint Lucia",
    "Saint Vincent and the Grenadines", "Samoa", "Sao Tome and Principe",
    "Saudi Arabia", "Senegal", "Serbia", "Seychelles", "Sierra Leone",
    "Singapore", "Slovakia", "Slovenia", "Solomon Islands", "Somalia",
    "South Africa", "South Sudan", "Spain", "Sri Lanka", "Sudan",
    "Sudan (until 2011)", "Suriname", "Sweden", "Switzerland",
    "Syrian Arab Republic", "Tajikistan", "Thailand", "The former Yugoslav Republic of Macedonia",
    "Timor-Leste", "Togo", "Tonga", "Trinidad and Tobago", "Tunisia",
    "Turkey", "Turkmenistan", "Uganda", "Ukraine", "United Arab Emirates",
    "United Kingdom of Great Britain and Northern Ireland", "United Republic of Tanzania",
    "United States of America", "Uruguay", "Uzbekistan", "Vanuatu",
    "Venezuela (Bolivarian Republic of)", "Viet Nam", "Yemen",
    "Zambia", "Zimbabwe", "Andorra", "Cook Islands", "Dominica",
    "Marshall Islands", "Nauru", "Niue", "Palau", "Saint Kitts and Nevis",
    "State of Palestine", "Tuvalu", "Monaco", "San Marino", "Germany, Federal Republic (former)",
    "India (until 1975)", "Kiribati (until 1984)", "South Viet Nam (former)")

years <- 2019:1962

merged_data = tibble(Location=countries) |> mutate(Period = list(years)) |> unnest(Period)

merged_data

```

```{r}

# pipelining each data set through data modifications it needs to become tidy
medical_doctors_data = medical_doctors_data |> replace_inidcator_to_tooltip()
birth_by_skilled_personel_data = birth_by_skilled_personel_data |> replace_inidcator_to_tooltip()
life_expectency_at_birth_data = life_expectency_at_birth_data |> replace_inidcator_to_tooltip() |> pivot_wider(names_from = Dim1, values_from=`Life expectancy at birth (years)`)
pharmacists_data = pharmacists_data |> replace_inidcator_to_tooltip()
infant_mortality_rate_data = infant_mortality_rate_data |> replace_inidcator_to_tooltip() |> pivot_wider(names_from = Dim1, values_from = `Infant mortality rate (probability of dying between birth and age 1 per 1000 live births)`) |> mutate(across(c(`Both sexes`, Male, Female), ~ as.numeric(sub(" .*", "", .x))))
maternal_mortality_ratio_data = maternal_mortality_ratio_data |> replace_inidcator_to_tooltip() |> mutate(across(`Maternal mortality ratio (per 100 000 live births)`, ~ as.numeric(sub(" .*", "", .x))))
incidence_of_malaria_data = incidence_of_malaria_data |> replace_inidcator_to_tooltip()
incidence_of_tuberculosis_data = incidence_of_tuberculosis_data |> replace_inidcator_to_tooltip() |> mutate(across(`Incidence of tuberculosis (per 100 000 population per year)`, ~ as.numeric(sub(" .*", "", .x))))

# air_pollution_death_rate = air_pollution_death_rate |> mutate(across(`First Tooltip`, ~ as.numeric(sub(" .*", "", .x)))) |> pivot_wider(names_from = c(Dim1, Dim2), values_from = `First Tooltip`) |> replace_inidcator_to_tooltip()

crude_sucide_rates = crude_sucide_rates |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
poison_mortality_rate = poison_mortality_rate |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
neo_natal_mortality_rate = neo_natal_mortality_rate |> mutate(across(`First Tooltip`, ~ as.numeric(sub(" .*", "", .x)))) |> replace_inidcator_to_tooltip() |> select(-c(Dim1))
new_hiv_infections = new_hiv_infections |> drop_na() |> mutate(across(`First Tooltip`, ~ as.numeric(sub(" .*", "", .x)))) |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
road_traffic_deaths = road_traffic_deaths |> replace_inidcator_to_tooltip()
under_5_mortality_rate = under_5_mortality_rate |> mutate(across(`First Tooltip`, ~ as.numeric(sub(" .*", "", .x)))) |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
number_of_dentists = number_of_dentists |> replace_inidcator_to_tooltip()
drinking_water_services = drinking_water_services |> replace_inidcator_to_tooltip()
# least_basic_sanitation_services = least_basic_sanitation_services |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
basic_hand_washing = basic_hand_washing |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
clean_fuel_and_teck = clean_fuel_and_teck |> replace_inidcator_to_tooltip()
tabacco_age_15 = tabacco_age_15  |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()
nursing_services = nursing_services |> replace_inidcator_to_tooltip()
sanitation_services = sanitation_services |> pivot_wider(names_from=Dim1, values_from=`First Tooltip`) |> replace_inidcator_to_tooltip()

```

```{r}

merged_data = full_join(merged_data, drinking_water_services, by=c("Location", "Period"))
merged_data = full_join(merged_data, number_of_dentists, by=c("Location", "Period"))
# merged_data = full_join(merged_data, least_basic_sanitation_services |> rename(sanitation_services=Total) |> select(-c(Urban, Rural)), by=c("Location", "Period"))
merged_data = full_join(merged_data, basic_hand_washing |> rename(basic_hand_washing_services=Total) |> select(-c(Urban, Rural)), by=c("Location", "Period"))
merged_data = full_join(merged_data, tabacco_age_15 |> rename(tabacco_age_15=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, nursing_services, by=c("Location", "Period"))
merged_data = full_join(merged_data, sanitation_services |> rename(total_sanitation_services=Total) |> select(-c(Urban, Rural)), by=c("Location", "Period"))
merged_data = full_join(merged_data, under_5_mortality_rate |> rename(under_5_mortality_rate=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, road_traffic_deaths |> mutate(`Estimated road traffic death rate (per 10 000 population)`=`Estimated road traffic death rate (per 100 000 population)`/10) |> select(-c(`Estimated road traffic death rate (per 100 000 population)`)), by=c("Location", "Period"))
merged_data = full_join(merged_data, new_hiv_infections |> rename(hiv_infections=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, neo_natal_mortality_rate |> mutate(`Neonatal mortality rate (per 10 000 live births)`=`Neonatal mortality rate (per 1000 live births)`*10) |> select(-c(`Neonatal mortality rate (per 1000 live births)`)), by=c("Location", "Period"))
merged_data = full_join(merged_data, poison_mortality_rate |> rename(poison_mortality_rate=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, crude_sucide_rates |> rename(sucide_rate=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, life_expectency_at_birth_data |> rename(life_expectency_at_birth=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, birth_by_skilled_personel_data, by=c("Location", "Period"))
merged_data = full_join(merged_data, medical_doctors_data, by=c("Location", "Period"))
merged_data = full_join(merged_data, pharmacists_data, by=c("Location", "Period"))
merged_data = full_join(merged_data, infant_mortality_rate_data |> rename(infant_mortality_rate=`Both sexes`) |> select(-c(Male, Female)), by=c("Location", "Period"))
merged_data = full_join(merged_data, maternal_mortality_ratio_data |> mutate(`Maternal mortality ratio (per 10 000 live births)`=`Maternal mortality ratio (per 100 000 live births)`*10) |> select(-c(`Maternal mortality ratio (per 100 000 live births)`)), by=c("Location", "Period"))
merged_data = full_join(merged_data, incidence_of_malaria_data |> mutate(`Malaria incidence (per 10 000 population at risk)`=`Malaria incidence (per 1 000 population at risk)`*10) |> select(-c(`Malaria incidence (per 1 000 population at risk)`)), by=c("Location", "Period"))
merged_data = full_join(merged_data, incidence_of_tuberculosis_data |> mutate(`Incidence of tuberculosis (per 10 000 population per year)`=`Incidence of tuberculosis (per 100 000 population per year)`*10) |> select(-c(`Incidence of tuberculosis (per 100 000 population per year)`)), by=c("Location", "Period"))

merged_data

```

```{r}

outcome_feat = c("Maternal mortality ratio (per 10 000 live births)", "Malaria incidence (per 10 000 population at risk)", "Incidence of tuberculosis (per 10 000 population per year)", "under_5_mortality_rate", "Estimated road traffic death rate (per 10 000 population)", "hiv_infections", "Neonatal mortality rate (per 10 000 live births)", "poison_mortality_rate", "sucide_rate", "life_expectency_at_birth", "infant_mortality_rate")

explanatory_feat = setdiff(colnames(merged_data), outcome_feat)
explanatory_feat = setdiff(explanatory_feat, c("Location", "Period"))

```

```{r}

non_na_per_period <- merged_data %>% select(-c(Location)) %>%
  rowwise() %>%
  mutate(non_na_count = sum(!is.na(c_across(where(~ !is.list(.x)))))) %>%
  ungroup() %>%
  group_by(Period) %>%
  summarise(total_non_na = sum(non_na_count))

ggplot(non_na_per_period, aes(x = Period, y = total_non_na)) +
  geom_col(fill = "seagreen") +
  labs(title = "Total Non-NA Values per Period",
       x = "Year",
       y = "Non-NA Value Count") +
  theme_minimal()

# lets keep the data only from 2000 on wards

merged_data = merged_data |> filter(Period >= 2000)

na_counts <- merged_data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "feature", values_to = "na_count")

ggplot(na_counts, aes(x = reorder(feature, na_count), y = na_count)) +
  geom_col(fill = "tomato") +
  coord_flip() +
  labs(title = "NA Values per Feature",
       x = "Feature",
       y = "NA Count") +
  theme_minimal()

valid_features <- merged_data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "feature", values_to = "na_count") %>%
  filter(na_count < 3000) %>%
  pull(feature)

cleaned_merged_data <- merged_data %>%
  select(all_of(valid_features))

```

```{r}

explanatory_feat = intersect(explanatory_feat, names(cleaned_merged_data))

```

```{r}

# get correlation between each pair of features and get the top 5 pairs

numeric_data <- cleaned_merged_data %>% 
  select(c(explanatory_feat)) %>%
  select(where(is.numeric))

cor_matrix <- cor(numeric_data, use = "pairwise.complete.obs")

cor_long <- cor_matrix %>%
  as_tibble(rownames = "feature1") %>%
  pivot_longer(-feature1, names_to = "feature2", values_to = "correlation") %>%
  filter(feature1 != feature2) %>%
  mutate(pair = map2_chr(feature1, feature2, ~ paste(sort(c(.x, .y)), collapse = "___"))) %>%
  distinct(pair, .keep_all = TRUE) %>%
  select(-pair)

top_5_corr <- cor_long %>%
  arrange(desc(abs(correlation))) %>%
  slice_head(n = 10)

```

```{r}

# remove most correlated outcome features

cleaned_merged_data = cleaned_merged_data |> select(-c(under_5_mortality_rate, `Neonatal mortality rate (per 10 000 live births)`, `Neonatal mortality rate (per 10 000 live births)`))
```

```{r, fig.width=10, fig.height=6}

# ggpairs(cleaned_merged_data |> select(-c(Location, Period)) |> drop_na(), progress = F)

```

```{r}

explanatory_feat_infant_mortality = c(explanatory_feat, "infant_mortality_rate")

merged_train_test_split = cleaned_merged_data |> select(explanatory_feat_infant_mortality) |> filter(!is.na(infant_mortality_rate)) |> initial_split(prop=0.8)

merged_data_train = training(merged_train_test_split)
merged_data_test = testing(merged_train_test_split)


```

```{r}

# recipes

base_recipe = recipe(merged_data_train) |> update_role(infant_mortality_rate, new_role = "outcome") |> update_role(all_of(explanatory_feat), new_role = "predictor") |> step_normalize(all_numeric_predictors())

recipe_imputed = recipe(merged_data_train) |> update_role(infant_mortality_rate, new_role = "outcome") |> update_role(all_of(explanatory_feat), new_role = "predictor") |> step_impute_mean(all_numeric_predictors()) |> step_normalize(all_numeric_predictors())

```

```{r}

# models

lin_reg <- linear_reg() |> set_engine("lm")
rf_model  <- rand_forest() |> set_engine("ranger", importance = "permutation") |> set_mode("regression")
xgb_model <- boost_tree()  |> set_engine("xgboost") |> set_mode("regression")
lasso_model <- linear_reg(penalty = tune(), mixture = 1) |> set_engine("glmnet") |> set_mode("regression")
knn_model <- nearest_neighbor(neighbors = tune(), weight_func = "rectangular") |> set_engine("kknn") |> set_mode("regression")

```

```{r}

# workflows

lr_workflow = workflow() |> add_model(lin_reg) |> add_recipe(recipe_imputed)
rf_workflow  <- workflow() |> add_model(rf_model) |> add_recipe(recipe_imputed)
xgb_workflow <- workflow() |> add_model(xgb_model) |> add_recipe(recipe_imputed)
knn_workflow <- workflow() |> add_model(knn_model) |> add_recipe(recipe_imputed)
lasso_workflow <- workflow() |> add_model(lasso_model) |> add_recipe(recipe_imputed)

```

```{r}
metrics <- metric_set(yardstick::rmse, yardstick::mae, yardstick::rsq)
```

```{r}
# cross validation for k

cv_folds <- vfold_cv(merged_data_train, v = 5)

knn_grid <- tibble(neighbors = seq(1, 10, by = 1))

knn_tune_results <- tune_grid(
  knn_workflow,
  resamples = cv_folds,
  grid = knn_grid,
  metrics = metrics
)

autoplot(knn_tune_results)

best_k <- select_best(knn_tune_results, metric = "rmse")

final_knn_workflow <- finalize_workflow(knn_workflow, best_k)

```

```{r}

cv_folds <- vfold_cv(merged_data_train, v = 5)

lasso_grid <- tibble(penalty = seq(0.01, 1, by=0.01))

lasso_tune_results <- tune_grid(
  lasso_workflow,
  resamples = cv_folds,
  grid = lasso_grid,
  metrics = metrics
)

autoplot(lasso_tune_results)

best_reg <- select_best(lasso_tune_results, metric = "rmse")

final_lasso_workflow <- finalize_workflow(lasso_workflow, best_reg)

```

```{r}

lr_fit  <- fit(lr_workflow, data = merged_data_train)
rf_fit  <- fit(rf_workflow, data = merged_data_train)
xgb_fit <- fit(xgb_workflow, data = merged_data_train)
knn_fit <- fit(final_knn_workflow, data = merged_data_train)
lasso_fit <- fit(final_lasso_workflow, data = merged_data_train)


lr_preds <- predict(lr_fit, merged_data_test) %>%
  bind_cols(merged_data_test %>% select(infant_mortality_rate)) %>%
  mutate(model = "Linear")

rf_preds <- predict(rf_fit, merged_data_test) %>%
  bind_cols(merged_data_test %>% select(infant_mortality_rate)) %>%
  mutate(model = "Random Forest")

xgb_preds <- predict(xgb_fit, merged_data_test) %>%
  bind_cols(merged_data_test %>% select(infant_mortality_rate)) %>%
  mutate(model = "XGBoost")

knn_preds <- predict(knn_fit, merged_data_test) %>%
  bind_cols(merged_data_test %>% select(infant_mortality_rate)) %>%
  mutate(model = "Knn")

lasso_preds <- predict(lasso_fit, merged_data_test) %>%
  bind_cols(merged_data_test %>% select(infant_mortality_rate)) %>%
  mutate(model = "Lasso")

all_preds <- bind_rows(lr_preds, rf_preds, xgb_preds, knn_preds, lasso_preds)

pred_tibble = all_preds %>%
  group_by(model) %>%
  metrics(truth = infant_mortality_rate, estimate = .pred)

# rmse: root mean square error
# rsq: R^2
# mae: mean absolute error
```

```{r}

pred_tibble |> ggplot() + geom_col(aes(x = model, y = .estimate, fill = .metric), position = position_dodge())

```
```{r, fig.width=10, fig.height=6}

lr_fit |> extract_fit_engine() |> check_model()

```


```{r}

# got explanation for linear regression 

tidy(lr_fit |> extract_fit_engine() |> summary()) |> ggplot() + geom_col(aes(x=reorder(term, abs(estimate)), y=abs(estimate))) + coord_flip()

```

```{r}

rf_fitted <- extract_fit_parsnip(rf_fit)

tidy(rf_fitted$fit$variable.importance) |> ggplot() + geom_col(aes(x=reorder(names, x), y=x)) + coord_flip()

```















